---
title: "Data Exploration DATA621 Final"
subtitle: "Modeling Housing Prices"
author: "Kai Lukowiak, Jaan Bernberg, Ilya Kats, Michael Muller"
date: "`r format(Sys.time(), '%B %d, %Y')`"


output: 
  pdf_document:
    df_print: kable
    toc: true
    toc_depth: 2
    fig_caption: yes
    
abstract: "This document provides an introduction to R Markdown, argues for its..."
keywords: "pandoc, r markdown, knitr"
bibliography: bibliography.bib
---
keywords: "pandoc, r markdown, knitr"

```{r options_pkgs, echo=F, warning=F, message=F, results=F}

##Use PDF for Final Paper
  # html_document:
  #   theme: yeti
  #   code_folding: hide
  #   toc: true
  #   toc_float:
  #     collapsed: true
  #     smooth_scroll: false

knitr::opts_chunk$set(
                      error = F
                      , message = F
                      #,tidy = T
                      , cache = T
                      , warning = T
                      , results = 'hide' #suppress code output
                      , echo = F #suppress code
                      , fig.show = 'hide' #suppress plots
                      )

install_load <- function(pkg){
  # Load packages & Install them if needed.
  # CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

# required packages
packages <- c("tidyverse","knitr", "ggthemes", "mice", "VIM", "RCurl", "knitcitations")

install_load(packages)

```



# 1. Introduction (Jaan)

## Data Set Origin

## Variables

## Cases

Describe the background and motivation of your problem.





```{r setup }
##Read data
url_train <- "https://raw.githubusercontent.com/kaiserxc/DATA621FinalProject/master/house-prices-advanced-regression-techniques/train.csv"
url_test <-  "https://raw.githubusercontent.com/kaiserxc/DATA621FinalProject/master/house-prices-advanced-regression-techniques/test.csv"

stand_read <- function(url){
  return(read.csv(text = getURL(url)))
}

o_train <- 
  stand_read(url_train) %>% 
  mutate(d_name = 'train')
o_test <- stand_read(url_test) %>% 
  mutate(SalePrice = NA, d_name = 'test')

full_set <- rbind(o_train, o_test)
# x <- plot_missing(full_set)
```



```{r data wrangle I}
na_review <- function(df){
  # returns df of vars w/ NA qty desc.
  na_qty <- colSums(is.na(df)) %>% as.data.frame(stringsAsFactors=F)
  colnames(na_qty) <- c("NA_qty")
  na_qty <- cbind('Variable' = rownames(na_qty), na_qty) %>% 
    select(Variable, NA_qty)
  rownames(na_qty) <- NULL
  
  na_qty <- na_qty %>% 
    arrange(desc(NA_qty)) %>% filter(NA_qty > 0) %>% 
    mutate(Variable = as.character(Variable)) %>% 
    mutate(Pct_of_Tot =  round(NA_qty/nrow(df), 4) * 100)
  
  return(na_qty)
}

first_pass <- full_set %>% 
  # first_pass is train.csv and test.csv combined for NA reviews 
  # and imputation planning and calculated columns
  mutate(House_Age_Yrs = YrSold - YearBuilt, 
         RemodAdd_Age_Yrs = YrSold - YearRemodAdd, 
         Garage_Age_Yrs = YrSold - GarageYrBlt) 
```



```{r na review}
naVars <- na_review(first_pass %>% select(-SalePrice))
naVars


set_aside <- c(2600, 2504, 2421, 2127, 2041, 2186, 2525, 1488, 949, 2349, 2218, 2219, 333)
#View(first_pass[is.na(first_pass$PoolQC), ]) # 2600, 2504, 2421
#View(first_pass[is.na(first_pass$GarageFinish), ]) # 2127
#View(first_pass[is.na(first_pass$GarageQual), ]) # 2127
#View(first_pass[is.na(first_pass$GarageCond), ]) # 2127
#View(first_pass[is.na(first_pass$BsmtCond), ]) # 2041, 2186, 2525
#View(first_pass[is.na(first_pass$BsmtExposure), ]) # 1488, 949, 2349
#View(first_pass[is.na(first_pass$BsmtQual), ]) # 2218, 2219
#View(first_pass[is.na(first_pass$BsmtFinType2), ]) # 333
#View(first_pass[is.na(first_pass$MasVnrType), ]) #

#qty
# first_pass[first_pass$PoolArea == 0, ]      # 2,906
# first_pass[is.na(first_pass$PoolQC), ]
# first_pass[is.na(first_pass$Alley), ]       # 2,721
# first_pass[is.na(first_pass$Fence), ]       # 2,348
# first_pass[first_pass$Fireplaces == 0, ]    # 1,420
# first_pass[is.na(first_pass$GarageType),]   # 157
# first_pass[is.na(first_pass$GarageArea),]   # 1
# first_pass[is.na(first_pass$GarageFinish),] # 159
# first_pass[first_pass$GarageArea == 0, ]    # 158
# first_pass[first_pass$TotalBsmtSF == 0, ]   # 79
# first_pass[is.na(first_pass$Electrical),]   # 1
```


```{r set aside problem cases}
set_asideA <- '2600|2504|2421|2127|2041|2186|2525|1488|949|2349|2218|2219|333' # 13
set_asideB <- '|2550|524|2296|2593' # negative values in '_Age' columns

x <- first_pass %>% 
  # exclude set_aside observations to fill in known NA's
  filter(!grepl(paste0(set_asideA, set_asideB), Id))
  
naVarsx <- na_review(x %>% select(-SalePrice))
naVarsx


nrow(x[x$PoolArea==0, ])   # 2,887
# x[is.na(x$MiscFeature),]   # 2,793
# x[is.na(x$Alley),]         # 2,700
# x[is.na(x$Fence),]         # 2,331
# x[is.na(x$FireplaceQu),]   # 1,414
# nrow(x[x$LotFrontage==0, ])# 486
# x[is.na(x$GarageArea),]    # 158
# x[x$TotalBsmtSF == 0, ]    # 78
```


```{r complete known na}
obtain_data <- function(df){
  # like first_pass but with imputation that addresses 
  # observations that have known NA's
  df %>%
    mutate(PoolQC = fct_explicit_na(PoolQC, na_level='NoP'),
           MiscFeature = fct_explicit_na(MiscFeature, na_level='NoM'),
           Alley = fct_explicit_na(Alley, na_level='NoA'),
           Fence = fct_explicit_na(Fence, na_level = 'NoF'),
           FireplaceQu = fct_explicit_na(FireplaceQu, na_level = 'NoFp'), 
           LotFrontage = ifelse(is.na(LotFrontage), 0, LotFrontage),
           
           # Note GarageYrBlt set to 9999 may be a problem
           GarageYrBlt = ifelse(is.na(GarageYrBlt), 9999, GarageYrBlt), 
           GarageFinish = fct_explicit_na(GarageFinish, na_level = 'NoG'), 
           GarageQual = fct_explicit_na(GarageQual, na_level = 'NoG'), 
           GarageCond = fct_explicit_na(GarageCond, na_level = 'NoG'), 
           # NOTE: Garage_Age_Yrs: 0 doesn't seem appropriate... 
           Garage_Age_Yrs = ifelse(is.na(Garage_Age_Yrs), 0, Garage_Age_Yrs),
           GarageType = fct_explicit_na(GarageType, na_level = 'NoG'), 
          
           BsmtQual = fct_explicit_na(BsmtQual, na_level = 'NoB'),
           BsmtCond = fct_explicit_na(BsmtCond, na_level = 'NoB'),
           BsmtExposure = fct_explicit_na(BsmtExposure, na_level = 'NoB'),
           BsmtFinType1 = fct_explicit_na(BsmtFinType1, na_level = 'NoB'),
           BsmtFinType2 = fct_explicit_na(BsmtFinType2, na_level = 'NoB')
           )
}
```

```{r recombine probs with updated}
probl_obs <- full_set %>% 
  mutate(House_Age_Yrs = YrSold - YearBuilt, 
         RemodAdd_Age_Yrs = YrSold - YearRemodAdd, 
         Garage_Age_Yrs = YrSold - GarageYrBlt) %>% 
  filter(grepl(paste0(set_asideA, set_asideB), Id))

known_obs <- full_set %>% 
  filter(!grepl(paste0(set_asideA, set_asideB), Id)) %>% 
  mutate(House_Age_Yrs = YrSold - YearBuilt, 
         RemodAdd_Age_Yrs = YrSold - YearRemodAdd, 
         Garage_Age_Yrs = YrSold - GarageYrBlt)

full_set_clean <- rbind(obtain_data(known_obs), probl_obs) %>% arrange(Id)
#View(full_set_clean)
summary(full_set_clean)
naVarsy <- na_review(full_set_clean %>% select(-SalePrice))
sum(naVarsy$NA_qty) # 176
# unique(full_set_clean$Alley) # NoA  Grvl Pave <NA>, levels: Grvl Pave NoA
# unique(full_set_clean$PoolQC) # NoP  Ex   <NA> Fa   Gd, levels: Ex Fa Gd NoP
# unique(full_set_clean$GarageYrBlt) # character!
```


```{r datatype review}
var_types <- function(df){
  # returns df of Variable name and Type from df
  var_df <- sapply(df, class) %>% as.data.frame()
  colnames(var_df) <- c("Var_Type")
  var_df <- cbind(var_df, 'Variable' = rownames(var_df)) %>% 
    select(Variable, Var_Type) %>% 
    mutate(Variable = as.character(Variable),Var_Type = as.character(Var_Type))
  return(var_df)
}

var_review <- var_types(full_set_clean %>% select(-c(Id,SalePrice,d_name)))

fac_vars <- var_review %>% filter(Var_Type == 'factor') %>% 
  select(Variable) %>% t() %>% as.character() # 43 total length(fac_vars)
num_vars <- var_review %>% filter(grepl('character|integer|numeric', Var_Type)) %>% 
  select(Variable) %>% t() %>% as.character() # 39 total but see GarageYrBlt #length(num_vars) 
```


Number of complete cases original: 0  
Number of complete cases after repairing known NA's: 2,861 ($\approx 98\%$)  
Number of true NA's: 58

```{r summary stats numeric}
sum(complete.cases(full_set %>% select(-SalePrice)))       # 0
sum(complete.cases(full_set_clean %>% select(-SalePrice))) # 2,861 ~ 98%
nrow(full_set_clean) - 2861 # 58 NA
stat_info <- psych::describe(full_set_clean %>% select(num_vars, -Id, -d_name))
stat_info[c(2:nrow(stat_info)),c(2:5,8:9,13:ncol(stat_info)-1)]
```

```{r summary stats factor}
summary(full_set_clean %>% select(fac_vars, -Id, -SalePrice, -d_name))

```

```{r}
train_data <- full_set_clean %>% filter(d_name == 'train') %>% select(-d_name)
test_data <- full_set_clean %>% filter(d_name == 'test') %>% select(-d_name)
##View(train_data)
dim(train_data)
dim(test_data)
```

```{r}
# Data Exploration Plots
#plot_boxplot()
```


```{r values with neg calc cols}
full_set_clean %>% 
  filter(Garage_Age_Yrs < 0 | RemodAdd_Age_Yrs < 0 | Garage_Age_Yrs < 0) # Ids c(524, 2296, 2550, 2593)
```

# 2. Data Exploration (Kai)






# 3. Data Preparation (Kyle)

"The ggplot2 package is used to plot images in layers [@wickham_ggplot2_2009]."
"You have to reference the "
@wickham_ggplot2_2009
@RCoreTeam
@Chang2015


```{r plot2.1}
hist(train_data$SalePrice, main = "Plot 3.1")
```


## Missing Value Imputation

### Literature review

### Methodology

## Variable Transformations

### Literature review

### Methodology

```{r unused mice}
# init = mice(first_pass, maxit=0) 
# meth = init$method
# predM = init$predictorMatrix
# 
# # The code below will remove the variable as a 
# # predictor but still will be imputed. Just for 
# # illustration purposes, I select the BMI 
# # variable to not be included as predictor during 
# # imputation.
# predM[, c('SalePrice')] = 0
# 
# # If you want to skip a variable from imputation 
# # use the code below. This variable will be 
# # used for prediction.
# meth[] = ""
# 
# # Now let specify the methods for imputing the 
# # missing values. There are specific methods 
# # for continues, binary and ordinal variables. 
# # I set different methods for each variable. 
# # You can add more than one variable in each method.
# 
# meth[c("BsmtExposure", "BsmtFinType2", "MasVnrType", 
#        "MasVnrArea", "Electrical")]="norm" 
# 
# 
# imputed = mice(clinsurf, method=meth, 
#                predictorMatrix=predM, m=5, 
#                printFlag = F)
# #Create a dataset after imputation.
# imputed <- complete(imputed)
# sapply(imputed, FUN = function(x) sum(is.na(x)))
# NA Qtys
# CAR_AGE 510
# HOME_VAL 464 
# YOJ 454
# INCOME 445
# AGE 6
```


# 4. Modeling (????)

### Literature review

### Methodology


# 5. Model Selection, Diagnostics & Conclusions (Ilya)

### Literature review

### Methodology

#Appendix


## Tables & Outputs

```{r Tables, ref.label=knitr::all_labels(), echo=F, eval=TRUE, results=T}

```

## Plots

```{r Plots, ref.label=knitr::all_labels(), echo=FALSE, fig.show="asis"}

```




## Code
```{r Code, ref.label=knitr::all_labels(),echo=TRUE , eval=FALSE}

```

## Bibliography

